---
title: "Build and deploy a stroke prediction model using R"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

**BISHAL DHUNGANA**

05-01-2024

**ABOUT DATA ANALYSIS REPORT**

This RMarkdown file contains the report of the data analysis done for the project on building and deploying a stroke prediction model in R. It contains analysis such as data exploration, summary statistics, and building the prediction models.

**DATA DESCRIPTION:**

According to the World Health Organization (WHO), STROKE IS THE 2ND LEADING CAUSE OF DEATH GLOBALLY, RESPONSIBLE FOR APPROXIMATELY 11% OF TOTAL DEATHS.

This dataset is used to predict whether a patient is likely to get stroke based on the input parameters like gender, age, various diseases, and smoking status. Each row in the data provides relevant information about the patient.


**TASK ONE: IMPORT DATA AND DATA PREPROCESSING**
Load data and install packages


```{r}
packages <- c("tidyverse", "lubridate", "ggplot2", "dplyr", "caret", "randomForest", "skimr", "gridExtra", "caTools", "corrplot", "ggcorrplot", "naniar")

# Install and load required packages
for (pkg in packages) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
}
```

```{r}
Data_Stroke <- read.csv('healthcare-dataset-stroke-data.csv')
summary(Data_Stroke)
```

```{r}
glimpse(Data_Stroke)
```

```{r}
skim(Data_Stroke)
```

```{r}
miss_scan_count(data = Data_Stroke, search = list("Unknown","N/A","Other"))
```

```{r}
##Convert NA to median in BMI
Data_Stroke$bmi <- as.numeric(Data_Stroke$bmi)
```

```{r}
idx <- complete.cases(Data_Stroke)
bmi_idx <- is.na(Data_Stroke$bmi)
median_bmi <- median(Data_Stroke$bmi, na.rm = TRUE)

Data_Stroke[bmi_idx,]$bmi <- median_bmi
colSums(is.na(Data_Stroke))
```

```{r}
##Check duplicates
sum(duplicated(Data_Stroke))
```

```{r}
colSums(Data_Stroke == 'N/A')
```

```{r}
colSums(Data_Stroke == '')
```

```{r}
Data_Stroke %>% count(gender)
```

```{r}
##Remove ID and filter out 'Other' values in Gender
Data_Stroke <- Data_Stroke %>% 
  select(-c(id)) %>% 
  filter(gender != "Other")
str(Data_Stroke)
```

```{r}
##Convert non-numeric variables to factors
Data_Stroke$stroke <- factor(Data_Stroke$stroke, levels = c(0,1), labels = c("No", "Yes"))
Data_Stroke$hypertension <- factor(Data_Stroke$hypertension, levels = c(0,1), labels = c("No", "Yes"))
Data_Stroke$heart_disease <- factor(Data_Stroke$heart_disease, levels = c(0,1), labels = c("No", "Yes"))
```



**TASK TWO: BUILD PREDICTION MODELS**
```{r}
d1 <- Data_Stroke %>%
  ggplot(aes(x = gender, fill = gender)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Gender Distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)
  
d2 <- Data_Stroke %>%
  ggplot(aes(x = hypertension, fill = hypertension)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Hypertenstion Distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)
  

d3 <- Data_Stroke %>%
  ggplot(aes(x = heart_disease, fill = heart_disease)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Heart Disease Distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)

d4 <- Data_Stroke %>%
  ggplot(aes(x = ever_married, fill = ever_married)) +
  geom_bar(fill = c("red","blue")) +
  ggtitle("Married distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)

d5 <- Data_Stroke %>%
  ggplot(aes(x = work_type, fill = work_type)) +
  geom_bar(fill = c("red", "blue","green","orange","aquamarine")) +
  ggtitle("Work type distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)

d6 <- Data_Stroke %>%
  ggplot(aes(x = stroke, fill = stroke)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Stroke distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)

d7 <- Data_Stroke %>%
  ggplot(aes(x = Residence_type, fill = Residence_type)) +
  geom_bar(fill = c("red", "blue")) +
  ggtitle("Residence distribution") +
  geom_text(aes(label=..count..), stat = "Count", vjust = 1.0)


grid.arrange(d1,d2,d3,d4,d5,d6,d7, ncol=2)
```

```{r}
Data_Stroke %>%
  ggplot(aes(x = gender, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet")) +
  ggtitle("Gender vs. Stroke") 
```

```{r}
Data_Stroke %>%
  ggplot(aes(x = hypertension, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet")) +
  ggtitle("Hypertension vs. Stroke")
```

```{r}
Data_Stroke %>%
  ggplot(aes(x = heart_disease, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet")) +
  ggtitle("Heart disease vs. Stroke") 
```

```{r}
Data_Stroke %>%
  ggplot(aes(x = Residence_type, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet")) +
  ggtitle("Residence type vs. Stroke")
```

```{r}
Data_Stroke %>%
  ggplot(aes(x = smoking_status, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet")) +
  ggtitle("Smoking status vs. Stroke")
```

```{r}
Data_Stroke %>%
  ggplot(aes(x = work_type, fill = stroke)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet"
                             )) +
  ggtitle("Type of Work vs. Stroke")
```

```{r}
Data_Stroke %>%
  ggplot(aes(x = avg_glucose_level, fill = stroke)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet"
  )) +
  ggtitle("Average Glucose level vs. Stroke")
```

```{r}
Data_Stroke %>% filter(between(bmi, 0, 60)) %>%
  ggplot(aes(x = bmi, fill = stroke)) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(values=c("aquamarine3",
                             "blueviolet"
  )) +
  ggtitle("Body Mass Index vs. Stroke")
```



**TASK THREE: EVALUATE AND SELECT PREDICTION MODELS**
```{r}
sample.split(Data_Stroke$stroke,SplitRatio = 0.8)->split_tag
train<-subset(Data_Stroke,split_tag==TRUE)
test<-subset(Data_Stroke,split_tag==FALSE)
dim(train)
dim(test)
```



**TASK FOUR: DEPLOY THE MODEL**
```{r}
set.seed(123)
rf <- randomForest(formula = stroke~.,data = train)
rf
```

```{r}
plot(rf)
```

```{r}
confusionMatrix(predict(rf,test),test$stroke)
```



**TASK FIVE: FINDINGS AND CONCLUSIONS**
As depicted above, our model boasts
an accuracy rate exceeding 95%, indicating that it underwent effective
training.
